...

### 2023-09-16 manuscript/01_main.Rmd
revised results

A total of `r sum_stat$DET.TOT.RAW %>% good_num()` detections of `r sum_stat$UNIQUE.SPEC.RAW %>% good_num()` bird species were made, of which `r round(100*(sum_stat$DET.TOT.RAW - sum_stat$DET.TOT.FILT)/sum_stat$DET.TOT.RAW)`% (`r (sum_stat$DET.TOT.RAW - sum_stat$DET.TOT.FILT) %>% good_num()`) did not meet the criteria for inclusion in the analysis. Of the `r sum_stat$DET.TOT.FILT %>% good_num()` detections of `r sum_stat$UNIQUE.SPEC.FILT %>% good_num()` species analysed, `r round(sum_stat$DET.AUD.FILT.PC)`% (`r sum_stat$DET.AUD.FILT %>% good_num()`) were purely auditory, 
`r round(sum_stat$DET.VIS.FILT.PC)`% (`r sum_stat$DET.VIS.FILT %>% good_num()`) were purely visual and `r round(sum_stat$DET.BOTH.FILT.PC)`% (`r sum_stat$DET.BOTH.FILT %>% good_num()`) were both, and this did not change majorly with weeks or across points *(Figure S3)*. The average number of detections per point count varied from `r sum_stat$DET.AVG.W01` in Week 1 to `r sum_stat$DET.AVG.W13` in Week 13 *(Figure 2)*. Among `r sum_stat$FORAGING.FILT %>% good_num()` records of birds foraging, just `r sum_stat$FORAGING.FILT.FRUIT %>% good_num()` were of frugivory.

At the community level, bird abundance was affected by observer, week, canopy heterogeneity, tree density, and dominant ground vegetation and its interaction with week. In spite of the significant correlations between canopy heterogeneity and canopy cover (`r show_cor(habvar$CC, habvar$CH, "pearson")`), and between canopy cover and tree density (`r show_cor(habvar$CC, habvar$TDens, "pearson")`), canopy heterogeneity and tree density remained uncorrelated (`r show_cor(habvar$CH, habvar$TDens, "pearson")`). Further, tree density also correlated negatively with proportion of deciduous trees at a significant level (`r show_cor(habvar$TPD, habvar$TDens, "pearson")`), but seemed to be the more important predictor in the models.

There was an overall increase in bird abundance towards the end of the study period in all sampling points---across the five categories of ground vegetation---but the magnitude of increase was starkly different in each. Points with *Rubus* sp. and those with *Vaccinium* sp. as the dominant ground vegetation had intermediate values of both average abundance and increase in abundance with time. Points with graminoids in the ground layer had the lowest overall abundance as well as the lowest increase *(Figure 3)*. Points with ground layers comprising moss and bare ground both had low abundances in the beginning but towards the end had the highest abundances, i.e., they showed the strongest increases. Bird abundance related positively and negatively with log-transformed variables of canopy heterogeneity and tree density respectively *(Figure 3)* and these trends did not change noticeably over the course of the study period.

Within the bird community, habitat selection was very different between invertivores and omnivores *(Table 3)*. Invertivore abundance was affected by observer, proportion of deciduous trees and dominant ground vegetation. Abundance was lower in sampling points with 'Graminoid' ground vegetation than points with other ground layers *(Figure 4)*. Abundance also decreased with deciduous proportion and, notably, since a polynomial term did not give a better fit, there is no evidence for the notion of mixed forests with intermediate deciduous proportions having higher diversity and abundance. Neither of the two trends changed detectably with time.

On the other hand, omnivore abundance was affected by observer, habitat class, tree density, and canopy heterogeneity, and it changed with time. Abundance of omnivores generally increased as the season progressed, and decreased with increasing tree density, but overall was higher in forest edges than in forest interiors or roadside patches (Figure 5). With canopy heterogeneity, the initial trend appeared to be slightly negative yet almost insignificant, but with every week the trend became more and more positive, such that in week 13 it was very strong. In other words, omnivores were initially indifferent to canopy heterogeneity, and as the season progressed their abundance did not change in relatively homogenous points, but greatly increased in highly heterogenous points (Figure 5).

@@

@@

@@

>>>>

We detected a total of `r sum_stat$DET.TOT.RAW %>% good_num()` birds of `r sum_stat$UNIQUE.SPEC.RAW %>% good_num()` species, of which `r round(100*(sum_stat$DET.TOT.RAW - sum_stat$DET.TOT.FILT)/sum_stat$DET.TOT.RAW)`% (`r (sum_stat$DET.TOT.RAW - sum_stat$DET.TOT.FILT) %>% good_num()`) did not meet the criteria for inclusion in the analysis. Of the `r sum_stat$DET.TOT.FILT %>% good_num()` detections of `r sum_stat$UNIQUE.SPEC.FILT %>% good_num()` species analysed, `r round(sum_stat$DET.AUD.FILT.PC)`% (`r sum_stat$DET.AUD.FILT %>% good_num()`) were purely auditory, 
`r round(sum_stat$DET.VIS.FILT.PC)`% (`r sum_stat$DET.VIS.FILT %>% good_num()`) were purely visual and `r round(sum_stat$DET.BOTH.FILT.PC)`% (`r sum_stat$DET.BOTH.FILT %>% good_num()`) were both, and this did not change majorly with weeks or across points *(Figure S3)*. The average number of detections per point count varied from `r sum_stat$DET.AVG.W01` in Week 1 to `r sum_stat$DET.AVG.W13` in Week 13 *(Figure 2)*. Among `r sum_stat$FORAGING.FILT %>% good_num()` records of birds foraging, just `r sum_stat$FORAGING.FILT.FRUIT %>% good_num()` were of frugivory.

Community-level bird abundance was affected by observer, week and dominant ground vegetation, as well as canopy heterogeneity, tree density, and their interaction. In spite of the significant correlations between canopy heterogeneity and canopy cover (`r show_cor(habvar$CC, habvar$CH, "pearson")`), and between canopy cover and tree density (`r show_cor(habvar$CC, habvar$TDens, "pearson")`), canopy heterogeneity and tree density remained uncorrelated (`r show_cor(habvar$CH, habvar$TDens, "pearson")`). Further, tree density also correlated negatively with proportion of deciduous trees at a significant level (`r show_cor(habvar$TPD, habvar$TDens, "pearson")`), but seemed to be the more important predictor in the models. *(Fig of correlations)*

Bird abundance was lowest in sampling points with graminoids as dominant ground vegetation, intermediate in those with *Rubus* sp., and similarly high across points with the other three categories (*Vaccinium* sp., moss, bare ground; *Figure 3)*. The overall abundance of the bird community increased over the course of the study period across all sampling points *(Figure 3)*. At medium to high levels of tree density, bird abundance increased with increasing canopy heterogeneity. At relatively low levels of canopy heterogeneity, abundance decreased with increasing tree density. The relationship between these three variables did not change noticeably over the season, but broke down at low levels of tree density and medium to high levels of canopy heterogeneity *(Figure 3)*.

Within the bird community, habitat selection was very different between invertivores and omnivores *(Table 3)*. Invertivore abundance was much lower in sampling points with graminoid ground vegetation *(Figure 4)* throughout the study period, but regardless of the dominant ground vegetation, points having moss in the ground layer witnessed an increase in invertivore abundance as the season progressed. On the other hand, omnivore abundance slightly decreased with increasing tree density, and was lower in interior forest patches as compared with roadside and edge patches *(Figure 5)*. While there was an overall increase in abundance as the season progressed, the increase was negligible in patches with low canopy heterogeneity and stark in those with high canopy heterogeneity *(Figure 5)*.

@@

negbin omn, so could allude to flocking

@@

In other words, omnivores were initially indifferent to canopy heterogeneity, and as the season progressed their abundance did not change in relatively homogenous points, but greatly increased in highly heterogenous points (Figure 5).

@@

TPD there is no evidence for the notion of mixed forests with intermediate deciduous proportions having higher diversity and abundance
